\sffamily \Large Supplementary Information\\[0.1cm] {\bfseries Introspection dynamics in asymmetric multiplayers games}

### 2023-03-02 latex/article.tex

%\title{\sffamily \Large Supplementary Information\\[0.1cm] {\bfseries Introspection dynamics in asymmetric multiplayers games}}
\title{\sffamily \Large {\bfseries Introspection dynamics in asymmetric multiplayers games}}
\date{\empty}
\author{\parbox[c]{16cm}{\centering \onehalfspacing \fontsize{11}{12}\selectfont Marta Couto$^{1*}$ and Saptarshi Pal$^{1*}$\\[0.2cm]
$^1$Max Planck Research Group Dynamics of Social Behavior, Max Planck Institute for Evolutionary Biology, 24306~Ploen, Germany}\\ \\
$^*$ \fontsize{11}{12}\selectfont co-first authors}

\begin{document}
\maketitle
\onehalfspacing
\section*{Abstract}

@@

\section*{Introduction}

@@

\section*{Model of introspection dynamics in multiplayer games}
We consider a normal form game with $N (\geq 2)$ players. In the game, a player, say player $i$, can play actions from their action set, $\A_i := \{a_{i,1}, a_{i,2}, ..., a_{i,m_i} \}$. The action set of player $i$ has $m_i$ actions. In this model, players only use pure strategies. Therefore, there are finitely many states of the game. More precisely, there are exactly $m_1 \times m_2 \times ... \times m_N$ states. We denote a state of the game by collecting the actions of all the players in the game in a vector, $\abf := (a_1, a_2, ..., a_N)$ where $\abf \in \A := \A_1 \times \A_2 \times ... \times \A_N$ and $a_i \in \A_i$. We also use the notation, $\abf := (a_i, \abf_{-i})$ to denote the state from the perspective of player $i$. In the state $(a_i, \abf_{-i})$, player $i$ plays the action $a_i \in \A_i$ and their co-players play the actions $\abf_{-i} \in \A_{-i}$ where $\A_{-i}$ is defined as $\A_{-i}:= \prod_{j \neq i} \A_j$. The payoff of a player depends on the state of the game. We denote the payoff of player $i$ in the state $\abf$ with $\pi_i(\abf)$ or $\pi_i(a_i, \abf_{-i})$. In this paper, we use bold font letters to denote vectors and matrices. We use the corresponding normal font letters with subscripts to denote elements of the vectors (or matrices). Since players only use pure strategies in this model, we use the terms strategies and actions interchangebly throughout the whole paper. \\ \\ 
\noindent In this model, players update their strategies over time using introspection dynamics \cite{Couto:NJP:2022}. At every time step, one randomly chosen player can update their strategy. The randomly chosen player, say $i$, currently playing  action $a_{i,k}$, compares their current payoff to the payoff that they would obtain if they played a randomly selected action $a_{i,l} \neq a_{i,k}$ from their action set $\A_i$. This comparison is done while assuming that the co-players do not change their respective actions. When the co-players of player $i$ play $\abf_{-i}$, player $i$ changes from action $a_{i,k}$ to the new action $a_{i,l}$ in the next round with probability, \\
\begin{equation}
 p_{a_{i,k} \to a_{i,l}} (\abf_{-i})= \frac{1}{1 + e^{\displaystyle -\beta(\pi_i(a_{i,l}, \abf_{-i}) - \pi_i(a_{i,k}, \abf_{-i}))}}
 \label{Eq:introspection-update}
\end{equation}
\\ \\ \noindent Here $\beta \in [0,\infty)$ is the selection strength parameter that represents the importance that players give to payoff differences while updating their actions. For $\beta = 0$, players update to a randomly chosen strategy with probablity $0.5$. For $\beta > 0$, players update to the alternative strategy under consideration with probablity greater than $0.5$ (or less than $0.5$) if the switch gives them an increase (or decrease) in the payoffs. \\ \\ 
\noindent Introspection dynamics can be studied by analyzing properties of the transition matrix, $\T$ of the resulting dynamical process. The transition matrix element $\mathrm{T}_{\abf,\bbf}$ denotes the conditional probability that the game goes to the state $\bbf$ in the next round if it is in state $\abf$ in the current round. In order to formally define the transition matrix, we first need to introduce some notations and definitions. We start by defining the neighbourhood set of $\abf$,

@@

\noindent The stationary distribution element $\mathrm{u}_\abf$ is the probability that state $\abf$ will be played by the players in the long run. Using the stationary distribution, one can calculate the marginal probabilities corresponding to each player's actions. That is, the probability that player $i$ plays action $a \in \A_i$ in the long run, $\xi_{i,a}$, can be computed as,
\begin{equation}
\mathbf{\xi}_{i,a} := \sum_{\qbf \in \A_{-i}} \mathrm{u}_{(a, \qbf)}
\label{Eq:marginal-definition}
\end{equation}
%\noindent Since the stationary distribution is a probability distribution, marginal distributions also have the same property. That is, for a player $i$, 
%\begin{equation}
%\sum_{k = 1}^{m_i} \xi_{a_{i,k}}= 1
%\label{Eq:marginal-prob-dist}
%\end{equation}
\section*{Additive games and their properties under introspection dynamics}
In this section we discuss the stationary properties of introspection dynamics when players learn to play strategies in a special class of games: additive games \cite{McAvoy:PlosCB:2015, Pena:JTB:2014}. In an additive game, the payoff difference that a player earns by making a unilateral switch in their actions is independent of what their co-players play. In other words, if none of the co-players change their current actions, the payoff difference earned by making a switch in actions is \emph{only} determined by the switch and not on the actions of the co-players'. Formally, in additive games, for any player $i$, any pair of actions $x,y \in \A_i$, and any $\qbf \in \A_{-i}$,
\begin{equation}
\pi_i(x, \qbf) - \pi_i(y, \qbf) =: f_i(x,y) 
\label{Eq:coplayer-independent-payoffdiff}
\end{equation}
\noindent is independent of $\qbf$ and only dependent on $x$ and $y$. In the literature, this property is sometimes called \emph{equal gains from switching} \cite{Pena:JTB:2014}. For games with this property, the stationary distribution of introspection dynamics takes a simple form,

@@

\subsection*{Example of an additive game: linear public goods game with 2 actions}

@@

\section*{Games with two actions and their properties under introspection dynamics}

@@

\subsection*{An example of a  game with two actions: the general public goods game}

@@

\section*{Application: Introspection learning in a game with cooperation and rewards}

@@

\section*{Discussion and conclusion}

@@

@@

\includegraphics[width =  \textwidth]{figures/figure1.eps}~\\[0.4cm]

@@

\includegraphics[width =  \textwidth]{figures/figure2.eps}~\\[0.4cm]

@@

\includegraphics[width =  0.65\textwidth, keepaspectratio]{figures/figure3.eps}~\\[0.4cm]

@@

\includegraphics[width =  \textwidth]{figures/figure4.eps}~\\[0.4cm]

>>>>

\title{\sffamily \Large Supplementary Information\\[0.1cm] {\bfseries Introspection dynamics in asymmetric multiplayers games}}
\title{\sffamily \Large {\bfseries Introspection dynamics in asymmetric multiplayers games}}
\date{\empty}
\author{\parbox[c]{16cm}{\centering \onehalfspacing \fontsize{11}{12}\selectfont Marta C. Couto$^{1*}$ and Saptarshi Pal$^{1*}$\\[0.2cm]
$^1$Max Planck Research Group Dynamics of Social Behavior, Max Planck Institute for Evolutionary Biology, 24306~Ploen, Germany}\\ \\
$^*$ \fontsize{11}{12}\selectfont co-first authors}

\begin{document}
\maketitle
\onehalfspacing
\section*{Author Information}
\textbf{First co-author:} Marta C. Couto \\
\noindent \textbf{Affiliation:} \\Max Planck Research Group Dynamics of Social Behavior, \\Max Planck Institute for Evolutionary Biology, 24306~Ploen, Germany \\ 
\noindent \textbf{Email:} mccouto@evolbio.mpg.de \\
\noindent \textbf{ORCID:} 0000-0002-2868-9685  \\ \\
\textbf{Second co-author:} Saptarshi Pal \\
\noindent \textbf{Affiliation:} \\Max Planck Research Group Dynamics of Social Behavior, \\Max Planck Institute for Evolutionary Biology, 24306~Ploen, Germany \\ 
\noindent \textbf{Email:} pal@evolbio.mpg.de \\
\noindent \textbf{ORCID:} 0000-0003-3805-1295
\newpage 
\section*{Abstract}

@@

\section{Introduction}

@@

\section{Model of introspection dynamics in multiplayer games}
We consider a normal form game with $N (\geq 2)$ players. In the game, a player, say player $i$, can play actions from their action set, $\A_i := \{a_{i,1}, a_{i,2}, ..., a_{i,m_i} \}$. The action set of player $i$ has $m_i$ actions. In this model, players only use pure strategies. Therefore, there are finitely many states of the game. More precisely, there are exactly $m_1 \times m_2 \times ... \times m_N$ states. We denote a state of the game by collecting the actions of all the players in the game in a vector, $\abf := (a_1, a_2, ..., a_N)$ where $\abf \in \A := \A_1 \times \A_2 \times ... \times \A_N$ and $a_i \in \A_i$. We also use the notation, $\abf := (a_i, \abf_{-i})$ to denote the state from the perspective of player $i$. In the state $(a_i, \abf_{-i})$, player $i$ plays the action $a_i \in \A_i$ and their co-players play the actions $\abf_{-i} \in \A_{-i}$ where $\A_{-i}$ is defined as $\A_{-i}:= \prod_{j \neq i} \A_j$. The payoff of a player depends on the state of the game. We denote the payoff of player $i$ in the state $\abf$ with $\pi_i(\abf)$ or $\pi_i(a_i, \abf_{-i})$. In this paper, we use bold font letters to denote vectors and matrices. We use the corresponding normal font letters with subscripts to denote elements of the vectors (or matrices). Since players only use pure strategies in this model, we use the terms strategies and actions interchangebly throughout the whole paper. \\ \\ 
\noindent In this model, players update their strategies over time using introspection dynamics \cite{Couto:NJP:2022}. At every time step, one randomly chosen player can update their strategy. The randomly chosen player, say $i$, currently playing  action $a_{i,k}$, compares their current payoff to the payoff that they would obtain if they played a randomly selected action $a_{i,l} \neq a_{i,k}$ from their action set $\A_i$. This comparison is done while assuming that the co-players do not change their respective actions. When the co-players of player $i$ play $\abf_{-i}$, player $i$ changes from action $a_{i,k}$ to the new action $a_{i,l}$ in the next round with probability, \\
\begin{equation}
 p_{a_{i,k} \to a_{i,l}} (\abf_{-i})= \frac{1}{1 + e^{\displaystyle -\beta(\pi_i(a_{i,l}, \abf_{-i}) - \pi_i(a_{i,k}, \abf_{-i}))}}
 \label{Eq:introspection-update}
\end{equation}
\\ \\ \noindent Here $\beta \in [0,\infty)$ is the selection strength parameter that represents the importance that players give to payoff differences while updating their actions. For $\beta = 0$, players update to a randomly chosen strategy with probablity $0.5$. For $\beta > 0$, players update to the alternative strategy under consideration with probablity greater than $0.5$ (or less than $0.5$) if the switch gives them an increase (or decrease) in the payoffs. \\ \\ 
\noindent Introspection dynamics can be studied by analyzing properties of the transition matrix, $\T$ of the resulting dynamical process. The transition matrix element $\mathrm{T}_{\abf,\bbf}$ denotes the conditional probability that the game goes to the state $\bbf$ in the next round if it is in state $\abf$ in the current round. In order to formally define the transition matrix, we first need to introduce some notations and definitions. We start by defining the neighbourhood set of $\abf$,

@@

\noindent The stationary distribution element $\mathrm{u}_\abf$ is the probability that state $\abf$ will be played by the players in the long run. Using the stationary distribution, one can calculate the marginal probabilities corresponding to each player's actions. That is, the probability that player $i$ plays action $a \in \A_i$ in the long run, $\xi_{i,a}$, can be computed as,
\begin{equation}
\mathbf{\xi}_{i,a} := \sum_{\qbf \in \A_{-i}} \mathrm{u}_{(a, \qbf)}
\label{Eq:marginal-definition}
\end{equation}
%\noindent Since the stationary distribution is a probability distribution, marginal distributions also have the same property. That is, for a player $i$, 
%\begin{equation}
%\sum_{k = 1}^{m_i} \xi_{a_{i,k}}= 1
%\label{Eq:marginal-prob-dist}
%\end{equation}
\section{Additive games and their properties under introspection dynamics}
In this section we discuss the stationary properties of introspection dynamics when players learn to play strategies in a special class of games: additive games \cite{McAvoy:PlosCB:2015, Pena:JTB:2014}. In an additive game, the payoff difference that a player earns by making a unilateral switch in their actions is independent of what their co-players play. In other words, if none of the co-players change their current actions, the payoff difference earned by making a switch in actions is \emph{only} determined by the switch and not on the actions of the co-players'. Formally, in additive games, for any player $i$, any pair of actions $x,y \in \A_i$, and any $\qbf \in \A_{-i}$,
\begin{equation}
\pi_i(x, \qbf) - \pi_i(y, \qbf) =: f_i(x,y) 
\label{Eq:coplayer-independent-payoffdiff}
\end{equation}
\noindent is independent of $\qbf$ and only dependent on $x$ and $y$. In the literature, this property is sometimes called \emph{equal gains from switching} \cite{Pena:JTB:2014}. For games with this property, the stationary distribution of introspection dynamics takes a simple form,

@@

\subsection{Example of an additive game: linear public goods game with 2 actions}

@@

\section{Games with two actions and their properties under introspection dynamics}

@@

\subsection{An example of a  game with two actions: the general public goods game}

@@

\section{Application: Introspection learning in a game with cooperation and rewards}

@@

\section{Discussion and conclusion}

@@

\section*{Statements and Declarations}
The authors declare that they have no competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

@@

\includegraphics[width =  \textwidth]{figures/Fig1.eps}~\\[0.4cm]

@@

\includegraphics[width =  \textwidth]{figures/Fig2.eps}~\\[0.4cm]

@@

\includegraphics[width =  0.65\textwidth, keepaspectratio]{figures/Fig3.eps}~\\[0.4cm]

@@

\includegraphics[width =  \textwidth]{figures/Fig4.eps}~\\[0.4cm]
